Exact asymptotic expressions for local clustering in hyperbolic random graphs

### 2018-11-02 introduction.tex
Minor update on main results and fixed proof of Theorem 1.1

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

@@

\begin{corollary}[Asymptotic behavior of local clustering function in $G_{\H,n}(\alpha,\nu)$]
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$ and $s_\alpha(k)$ defined as in~\eqref{eq:def_scaling_function}. Then, as $n \to \infty$,
\[
	\frac{c_{\H,n}(k_n)}{s_\alpha(k_n)} \, \stackrel{L^1}{\rightarrow} \, 
	\begin{cases}
				C_\alpha \left(\frac{4\alpha \nu}{\pi\left(2\alpha - 1\right)}\right)^{4\alpha - 2}
				&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
				\frac{3\nu}{\pi} &\mbox{if } \alpha = \frac{3}{4},\\
				\frac{8\alpha \nu}{\pi\left(4\alpha - 3\right)} &\mbox{if } \alpha \ge \frac{3}{4},
		\end{cases}
\]
where $\stackrel{L^1}{\to}$ denotes converges in expectation.
\end{corollary}

@@

@@

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma~\ref{lem:N_k_HP_P}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n^{-(2\alpha + 1)} n$ tends to zero as $n \to \infty$ we have that 
\[
	\lim_{n \to \infty} \Exp{N_{\H,n}(k_n)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n$ is not $\smallO{n^{\frac{1}{2\alpha + 1}}}$.

>>>>

\begin{theorem}[Limit for local clustering function in $G_{\H,n}(\alpha, \nu)$]
\label{thm:local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$ and $(k_n)_{n \ge 1}$ be any positive sequence, possibly constant, such that $k_n \ge 3$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then, 
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{c_\infty(k_n)} - 1\right|} = 0,
\]
where the function $c_\infty(k)$ is defined for $\alpha \ne 1$ as
\begin{align*}
c_\infty(k)  =&\frac{1}{8\alpha (\alpha-1)\Gamma^+(k-2\alpha,\xi)} \left( -\Gamma^+(k - 2 \alpha, \xi) - 2\frac{\alpha (\alpha - 1/2)^2 \xi^{2} \Gamma^+(k - 2 \alpha - 2, \xi)}{(\alpha - 1)} \right. \\ 
&\left.+ 8 \alpha (\alpha - 1/2) \xi \Gamma^+(k - 2 \alpha - 1,\xi) \right.\\ 
&\left.+ 4\xi^{4\alpha - 2} \Gamma^+(k - 6 \alpha + 2, 
      \xi) \left( \frac{2^{ - 4\alpha}(3 \alpha - 1)}{(\alpha - 1)} + (\alpha - 1/2) B^-(1/2; 1 + 2 \alpha, -2 + 2 \alpha) \right)  \right.\\ 
&\left.+ \xi^{k-2\alpha} \Gamma(2\alpha+1)e^{-\xi} U(2\alpha+1,1+k-2\alpha,\xi) \right. \\ 
&\left.- \xi^{4\alpha-2} \Gamma(2\alpha+1)\MeijerG{3}{0}{2}{3}{1,3-2\alpha}{3-4\alpha,-6\alpha+k+2,0}{\xi}  \right)
\end{align*}
and for $\alpha = 1$ as
\begin{align*}
	c_\infty(k) &= \frac{9 \eta^3}{2 k!} 	
		\Gamma^+(k-3,\eta)-\frac{\xi_{\alpha,\nu}^4}{k!}\frac{7+\pi^2}{4}\Gamma^+(k-4,\eta)\\
	&\hspace{10pt}+ \frac{\eta^k}{2k!}\int_0^1 (1-4z+3z^2)\ln(1-z)z^{1-k}e^{-\eta/z}\dd z\\ 
	&\hspace{10pt}+ \frac{\eta^k}{k!}\int_0^1 z^{3-k} \Li_2(z) e^{-\eta/z} \dd z,
\end{align*}
with $\eta = 4\nu/\pi$ and $\Li_2(z) = \sum_{t = 1}^\infty z^t/t^2$, the dilogarithm function.
\end{theorem}

@@

\begin{corollary}[Asymptotic behavior of local clustering function in $G_{\H,n}(\alpha,\nu)$]
\label{cor:asymptotic_local_clustering_hyperbolic}
Let $\alpha > \frac{1}{2}$, $\nu > 0$, $(k_n)_{n \ge 1}$ be any positive sequence such that $k_n \to \infty$ and $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$ and $s_\alpha(k)$ defined as in~\eqref{eq:def_scaling_function}. Then, as $n \to \infty$,
\[
	\frac{c_{\H,n}(k_n)}{s_\alpha(k_n)} \, \stackrel{L^1}{\rightarrow} \, 
	\begin{cases}
				C_\alpha \left(\frac{4\alpha \nu}{\pi\left(2\alpha - 1\right)}\right)^{4\alpha - 2}
				&\mbox{if } \frac{1}{2} < \alpha < \frac{3}{4},\\
				\frac{3\nu}{\pi} &\mbox{if } \alpha = \frac{3}{4},\\
				\frac{8\alpha \nu}{\pi\left(4\alpha - 3\right)} &\mbox{if } \alpha \ge \frac{3}{4},
		\end{cases}
\]
where $\stackrel{L^1}{\to}$ denotes converges in expectation.
\end{corollary}

@@

\paragraph{Uniform convergence}

Our results for the local clustering function in particular imply uniform convergence of $c_{\H,n}(k)$ for all $3 \le k \le a_n$ where $a_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. To see this let
\[
	b_n = \arg \max_{3 \le k \le a_n} \Exp{\left|\frac{c_{\H ,n}(k)}{c_\infty(k)} - 1\right|}.
\]
Then $b_n \le a_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$ and therefore by Theorem~\ref{thm:local_clustering_hyperbolic}
\[
	\lim_{n \to \infty} \max_{3 \le k \le a_n} \Exp{\left|\frac{c_{\H ,n}(k)}{c_\infty(k)} - 1\right|} 
	= \lim_{n \to \infty}\Exp{\left|\frac{c_{\H ,n}(b_n)}{c_\infty(b_n)} - 1\right|} = 0.
\]
The same is true for the result in Corollary~\ref{cor:asymptotic_local_clustering_hyperbolic}.

\paragraph{Maximum scaling for $k_n$ is $n^{\frac{1}{2\alpha +1}}$}
All our results for clustering in the hyperbolic random graph are valid for any sequence $k_n$ such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the optimal scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_local_clustering_general} we have that $c_{\H,n}(k_n) = 0$ if $N_{\H,n}(k_n) = 0$. Hence, it follows by Markov's inequality that for any positive function $f$
\begin{align*}
	\Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	&\ge \Prob{N_{\H,n}(k_n) = 0} \ge 1 - \Exp{N_{\H,n}(k_n)}.
\end{align*}
We shall later establish (see Lemma~\ref{lem:N_k_H_P}) that $\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}$. Therefore if $k_n$ is such that $k_n^{-(2\alpha + 1)} n$ tends to zero as $n \to \infty$ we have that 
\[
	\lim_{n \to \infty} \Exp{N_{\H,n}(k_n)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c_{\H,n}(k_n)}{f(k_n)} - 1\right|} 
	\ge \lim_{n \to \infty} 1 - \Exp{N_{\H,n}(k_n)} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:clustering_coefficient_hyperbolic} to hold as soon as $k_n$ is not $\smallO{n^{\frac{1}{2\alpha + 1}}}$.

@@

### 2018-11-02 proof_overview.tex
Minor update on main results and fixed proof of Theorem 1.1

We can now use this result to prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}.

@@

\PvdH{Major errors. FIX!}

First, let $a_n$ be any sequence such that $a_n \to \infty$. Then
\begin{align*}
	c_{\H,n} &= \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) \le a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}}
		+ \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) > a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}}\\
	&= \frac{1}{n} \sum_{t = 3}^{a_n} \sum_{i = 1}^n \ind{D_{\H,n}(i) = t} \frac{T_{\H,n}(i)}{\binom{t}{2}}
		+ \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) > a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}}\\
	&= \frac{1}{n} \sum_{t = 3}^{a_n} \Exp{N_{\H,n}(t)} c_{\H,n}(t)
		+ \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) > a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}},
\end{align*}
where the expectation of last term is bounded from above by $\Prob{D_{\H,n} > a_n}$. We have a similar result for $c_\infty$. Therefore
\begin{align*}
	\Exp{|c_{\H,n} - c_\infty|}
	&\le \Exp{\left|\frac{1}{n} \sum_{t = 3}^{a_n} \left(\Exp{N_{\H,n}(t)} c_{\H,n}(t) - 	
		\Exp{N_\Pcal(t)}c_{\infty}(t)\right)\right|}\\ 
	&\hspace{10pt}+ \Prob{D_{\H,n} > a_n} + \Prob{D_{\Pcal} > a_n},
\end{align*}
and since the last to probabilities are $\smallO{1}$ it is enough to prove that
\[
	\lim_{n \to \infty} \Exp{\left|\frac{1}{n} \sum_{t = 3}^{a_n} \left(\Exp{N_{\H,n}(t)} c_{\H,n}(t) - 	
			\Exp{N_\Pcal(t)}c_{\infty}(t)\right)\right|} = 0.
\]

We write
\begin{align*}
	\Exp{\left|\frac{1}{n} \sum_{t = 3}^{a_n} \left(\Exp{N_{\H,n}(t)} c_{\H,n}(t) - 	
		\Exp{N_\Pcal(t)}c_{\infty}(t)\right)\right|}
	&\le \frac{1}{n} \sum_{t = 3}^{a_n} \Exp{N_\Pcal(t)} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}\\
	&\hspace{10pt}+ \frac{1}{n} \sum_{t = 3}^{a_n} \Exp{\left|N_{\H,n}(t) - N_\Pcal(t)\right|}\Exp{c_\infty(t)},
\end{align*}
and will show that both terms go to zero for some appropriately chosen sequence $a_n$.

Note that Theorem~\ref{thm:local_clustering_hyperbolic} states that for any sequence $k_n  = \smallO{n^{1/(2\alpha + 1)}}$
\[
	\lim_{n \to \infty} \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} = 0.
\]

Let $b_n := \log(n)^{-1} n^{1/(2\alpha + 1)}$ and define the sequence $k_n$ by
\[
	k_n = \arg \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}.
\]
Then $k_n = \smallO{n^{1/(2\alpha + 1)}}$ and by Theorem~\ref{thm:local_clustering_hyperbolic}
\[
	\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} = \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} \to 0.
\]
Similarly we note that
\[
	\arg \max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} \to 0.
\]

Now define
\[
	a_n := \left\lfloor \min \left\{
	\left(\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}\right)^{-1/2},
	\left(\arg \max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|}\right)^{-1/2},
	b_n - 1 \right\} \right\rfloor
\]
so that $a_n < b_n = \smallO{n^{1/(2\alpha + 1)}}$, $a_n \to \infty$, 
\[
	a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} 
	\le a_n \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} \to 0,
\]
and similarly
\[
	a_n \max_{3 \le t \le a_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} \to 0.
\]

Then, since $\Exp{N_\Pcal(t)} = \bigT{1} t^{-(2\alpha + 1)}$
\begin{align*}
	&\hspace{-30pt}\frac{1}{n} \sum_{t = 3}^{a_n} \Exp{N_\Pcal(t)} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}\\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}
	= \smallO{1},
\end{align*}
while $\Exp{c_\infty(t)} \le 1$ implies that
\begin{align*}
	&\hspace{-30pt}\frac{1}{n} \sum_{t = 3}^{a_n} \Exp{\left|N_{\H,n}(t) - N_\Pcal(t)\right|}\Exp{c_\infty(t)} \\
	&\le \frac{1}{n} \sum_{t = 3}^{a_n} \Exp{N_\Pcal(t)} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} \\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|} 
		= \smallO{1}.
\end{align*}
This finishes the proof.
\end{proof}

>>>>

We can now use this result to prove Theorem~\ref{thm:clustering_coefficient_hyperbolic}. Before we state the proof we summarize some results regarding the degrees in the hyperbolic model. The proof of this lemma follows directly from Lemma~\ref{lem:diff_Nk_hyperbolic_binomial_poisson} and Lemma~\ref{lem:N_k_HP_P}, see Section~\ref{sec:coupling_H_P_n}.

\begin{lemma}\label{lem:N_k_H_P}
Let $\alpha > \frac{1}{2}$ and $k_n$ be a positive sequence, possibly constant, such that $k_n = \smallO{n^{\frac{1}{2\alpha + 1}}}$. Then
\[
	\lim_{n\to \infty} \Exp{\left|\frac{\Exp{N_{\H,n}(k_n)}}{n \Prob{D_\Pcal = k_n}} - 1\right|} = 0
\]
and hence, as $n \to \infty$
\[
	\Exp{N_{\H,n}(k_n)} = \bigT{n k_n^{-(2\alpha + 1)}}.
\]
\end{lemma}

@@

First, let $a_n$ be any sequence such that $a_n \to \infty$. Then
\begin{align*}
	c_{\H,n} &= \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) \le a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}}
		+ \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) > a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}}\\
	&= \frac{1}{n} \sum_{t = 3}^{a_n} \sum_{i = 1}^n \ind{D_{\H,n}(i) = t} \frac{T_{\H,n}(i)}{\binom{t}{2}}
		+ \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) > a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}}\\
	&= \frac{1}{n} \sum_{t = 3}^{a_n} \Exp{N_{\H,n}(t)} c_{\H,n}^\ast(t)
		+ \frac{1}{n} \sum_{i = 1}^n \ind{D_{\H,n}(i) > a_n} \frac{T_{\H,n}(i)}{\binom{D_{\H,n}(i)}{2}},
\end{align*}
where the expectation of last term is bounded from above by $\Prob{D_{\H,n} > a_n}$. Similarly,
\[
	c_\Pcal = \sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} c_\Pcal(t) + \sum_{t > a_n} \Prob{D_\Pcal = t} c_\Pcal(t)
\]
Recall that (see [??]) $c_\infty = c_\Pcal$ and $c_\infty(k) = c_\Pcal(k)$. Therefore
\begin{align*}
	\Exp{|c_{\H,n} - c_\Pcal|}
	&\le \Exp{\left|\frac{1}{n} \sum_{t = 3}^{a_n} \left(\Exp{N_{\H,n}(t)} c_{\H,n}^\ast(t) - 	
		n \Prob{D_\Pcal = t}c_\Pcal(t)\right)\right|}\\ 
	&\hspace{10pt}+ \Prob{D_{\H,n} > a_n} + \Prob{D_{\Pcal} > a_n},
\end{align*}
and since the last to probabilities are $\smallO{1}$ it is enough to prove that
\[
	\lim_{n \to \infty} \Exp{\left|\frac{1}{n} \sum_{t = 3}^{a_n} \left(\Exp{N_{\H,n}(t)} c_{\H,n}^\ast(t) - 	
			n \Prob{D_\Pcal = t}c_\Pcal(t)\right)\right|} = 0.
\]

We write
\begin{align*}
	&\hspace{-30pt}\Exp{\left|\frac{1}{n} \sum_{t = 3}^{a_n} \left(\Exp{N_{\H,n}(t)} c_{\H,n}^\ast(t) - 	
		n \Prob{D_\Pcal = t}c_\Pcal(t)\right)\right|}\\
	&\le \sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|c_{\H,n}^\ast(t) - c_{\infty}(t)\right|}\\
	&\hspace{10pt}+ \frac{1}{n} \sum_{t = 3}^{a_n} \Exp{\left|N_{\H,n}(t) - n \Prob{D_\Pcal = t}\right|}\Exp{c_\Pcal(t)},
\end{align*}
and will show that both terms go to zero for some appropriately chosen sequence $a_n$.

Note that Theorem~\ref{thm:local_clustering_hyperbolic} and Lemma~\ref{lem:clustering_ast_H} together state that for any sequence $k_n  = \smallO{n^{1/(2\alpha + 1)}}$
\begin{equation}\label{eq:convergence_c_H_P}
	\lim_{n \to \infty} \Exp{\left|c_{\H,n}(k_n) - c_{\Pcal}(k_n)\right|} = 0.
\end{equation}

Let $b_n := \log(n)^{-1} n^{1/(2\alpha + 1)}$ and define the sequence $k_n$ by
\[
	k_n = \arg \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\Pcal}(t)\right|}.
\]
Then $k_n = \smallO{n^{1/(2\alpha + 1)}}$ and by~\eqref{eq:convergence_c_H_P}
\[
	\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\Pcal}(t)\right|} = \Exp{\left|c_{\H,n}(k_n) - c_{\infty}(k_n)\right|} \to 0.
\]
Similarly we note that by Lemma~\ref{lem:N_k_H_P}
\[
	\max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} \to 0.
\]

Now define
\[
	a_n := \left\lfloor \min \left\{
	\left(\max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|}\right)^{-1/2},
	\left(\arg \max_{3 \le t \le b_n} \Exp{\left|\frac{N_{\H,n}(t)}{N_\Pcal(t)} - 1\right|}\right)^{-1/2},
	b_n - 1 \right\} \right\rfloor
\]
so that $a_n < b_n = \smallO{n^{1/(2\alpha + 1)}}$, $a_n \to \infty$, 
\[
	a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} 
	\le a_n \max_{3 \le t \le b_n} \Exp{\left|c_{\H,n}(t) - c_{\infty}(t)\right|} \to 0,
\]
and similarly
\[
	a_n \max_{3 \le t \le a_n} \Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} \to 0.
\]

Then, since $\Exp{\Prob{D_\Pcal = t}} = \bigT{1} t^{-(2\alpha + 1)}$
\begin{align*}
	&\hspace{-30pt}\sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|c_{\H,n}^\ast(t) - c_{\Pcal}(t)\right|}\\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} \Exp{\left|c_{\H,n}^\ast(t) - c_{\Pcal}(t)\right|}
	= \smallO{1},
\end{align*}
while $\Exp{c_\Pcal(t)} \le 1$ implies that
\begin{align*}
	&\hspace{-30pt}\frac{1}{n} \sum_{t = 3}^{a_n} \Exp{\left|N_{\H,n}(t) - n \Prob{D_\Pcal = t}\right|}\Exp{c_\Pcal(t)} \\
	&\le \sum_{t = 3}^{a_n} \Prob{D_\Pcal = t} \Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} \\
	&\le \bigT{1} 3^{-(2\alpha+1)} a_n \max_{3 \le t \le a_n} 
		\Exp{\left|\frac{N_{\H,n}(t)}{n \Prob{D_\Pcal = t}} - 1\right|} = \smallO{1}.
\end{align*}
This finishes the proof.
\end{proof}
